---
title: "Vignette for the test package from the Building R Packages course in Coursera"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{testpackage-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(testpackagedevtools)
```

# Overview

This package provides functions for working with accident data stored in compressed CSV files. The main tasks covered by the package are:

- generating file names from years
- reading a single accident file, corresponding to a single year
- reading several accident files of different years at once
- summarizing accident counts by month and year
- plotting accidents for a given state on a map

# Create a file name from a year

The function `make_filename()` converts a year into the expected compressed file name format.

```{r}
make_filename(2013)
make_filename("2015")
```

This is useful when the data files follow a consistent naming convention such as the following included example files:

```r
accident_2013.csv.bz2
accident_2014.csv.bz2
accident_2015.csv.bz2
```

# Read a single accident file

The function `fars_read()` reads one local compressed CSV file and returns a tibble.

```{r, eval = FALSE}
dat <- fars_read("accident_2013.csv.bz2")
dat
```

The function checks whether the file exists before attempting to read it. If the file is missing, it throws an error.

```{r, eval = FALSE}
fars_read("does_not_exist.csv.bz2")
```

# Read several years at once

The function `fars_read_years()` takes a vector of years and tries to read the corresponding files. For each valid file, it returns a tibble containing only the `MONTH` column and an additional `year` column.

```{r, eval = FALSE}
year_list <- fars_read_years(c("2013", "2014", "2015"))
year_list
```

The output is a list, with one element per requested year.

For missing or invalid years, the function returns `NULL` for that element and issues a warning instead of stopping completely.

```{r, eval = FALSE}
year_list <- fars_read_years(c("2013", "2014", "2012"))
year_list
```

This makes it possible to process a vector of years even if some files are not available.

# Summarize accident counts by month and year

The function `fars_summarize_years()` combines the yearly data and counts how many accidents are recorded for each month in each year.

```{r, eval = FALSE}
summary_tbl <- fars_summarize_years(c("2013", "2014", "2015"))
summary_tbl
```

The result is a tibble where:

- each row represents a month
- each year becomes a separate column
- the values show the number of accidents for that month and year

This is useful for comparing seasonal patterns across years.

# Plot accidents for one state

The function `fars_map_state()` plots the accident locations for a selected year and U.S. state.

```{r, eval = FALSE}
fars_map_state(state.num = 1, year = 2013)
```

The `state.num` argument must match a valid state code present in the data. If the state code is invalid, the function stops with an error.

```{r, eval = FALSE}
fars_map_state(state.num = 100, year = 2013)
```

The function also handles invalid coordinates by converting out-of-range latitude and longitude values to missing values before plotting.

# A complete example workflow

The following example shows how the package functions can be used together.

```{r, eval = FALSE}
file_2013 <- make_filename(2013)

dat_2013 <- fars_read(file_2013)

year_data <- fars_read_years(c(2013, 2014, 2015))

year_summary <- fars_summarize_years(c(2013, 2014, 2015))

fars_map_state(state.num = 1, year = 2013)
```
